Draft_Application_Stochastic_Volatility

### 2021-07-20 Real Dataset SV.Rmd
commit

In order to evaluate the different filters according to the evaluation tool represented by the realized volatility, we start once more from the visual inspection of the plots. For any of the three indices, below graphs are shown including both the filtered volatility obtained from the filtered states (for any of the Sequential MC methods described in the theoretical section) and the corresponding rescaled realised volatility series.\footnote{For simplicity,troughout the rest of this section we will omit "rescaled" when refering to the rescaled realised volatility.} With respect to the visual analysis we conducted in section \textit{fill with the number of the section}, we can now use such a proxy as to determine the relative performance of the competing filters.\

Once more, most algorithms used (aside from SIS and Liu-West) seem to provide indistinguishable results and, generally speaking, to underestimate the volatility (as indicated by the realised volatility proxy). The SIS algorithm appears completely inadequate to describe the "true" volatility: as we argued in section XX, the degeneracy of its weights on the particles prevents it from delivering a good enough approximation of the target distribution (\textit{check whether this is the correct intuition}. On top of that, it does not seem the case that the SIS algorithm, by falling short of the SV target distribution, ends up fitting the proxy better.\footnote{As we will now see, this seems instead to be the case for our preferred algorithms, namely those adopting adaptive resampling rules: a better approximation of the SV target filtering distribution yields slightly higher RMSEs and MAEs than those resulting from less accurate MC approximations.}\
On the other hand, the Liu and West filter offers the best performance among all filters, at least visually. This suggests that, since the Liu and West particle filter is built on a different underlying model than the other filters considered, this model may offer a better representation of the data analyzed. Once more, these considerations hold for all of the three indices.\

@@

Finally, we can see from Table XX the RMSE and MAE obtained with a number of particles equal to 50000 for the SP500 index and the EuroStoxx50. In this case, although the number of particles increases, the RMSE and MAE increase (although by very small amounts). This is evidence supporting the idea that the model we have selected underlying for the first two filters is inappropriate for describing this specific data. Indeed, as widely discussed in SECTION XX, it is obsolete and surpassed by more advanced and precise models.

@@

>>>>

In order to evaluate the different filters according to the evaluation tool represented by the realized volatility, we start once more from the visual inspection of the plots. For any of the three indices, below graphs are shown including both the filtered volatility obtained from the filtered states (for any of the Sequential MC methods described in the theoretical section) and the corresponding rescaled realised volatility series.\footnote{For simplicity,troughout the rest of this section we will omit "rescaled" when refering to the rescaled realised volatility.} As compared to the visual analysis we conducted in section \textit{fill with the number of the section}, we can now use such a proxy to determine the relative performance of the competing filters.\

Once more, most algorithms used (aside from SIS and Liu-West) seem to provide indistinguishable results and, generally speaking, to underestimate the volatility (as indicated by the realised volatility proxy). The SIS algorithm appears completely inadequate to describe the "true" volatility process: as we argued in section XX, the degeneracy of its weights on the particles prevents it from delivering a good enough approximation of the target distribution (\textit{check whether this is the correct intuition}. On top of that, it does not seem the case that the SIS algorithm, by falling short of the SV target distribution, ends up fitting the proxy better.\footnote{As we will now see, this seems instead to be the case for our preferred algorithms, namely those adopting adaptive resampling rules: a better approximation of the SV target filtering distribution yields slightly higher RMSEs and MAEs than those resulting from less accurate MC approximations.}\
On the other hand, the Liu and West filter offers the best performance among all filters, at least visually. This suggests that, since the Liu and West particle filter is built on a different underlying model than the other filters considered, this model may offer a better representation of the data analyzed. Once more, these considerations hold for all of the three indices.\

@@

Finally, we can see from Table XX the RMSE and MAE obtained with a number of particles equal to 50000, for the SP500 and the EuroStoxx50 indices. We consider the Bootstrap particle filter and the Auxiliary particle filter, which we employed in section XX; in addition to these, we add the Liu and West filter, which did not feature in that section of this work.\
In the case of the SP500 index, although the number of particles increases, the RMSE and MAE increase (although by very small amounts) for both the particle filters based on our usual specification of the SV model. Instead, for the Liu and West filter we see some improvements, although the small extent suggests that also the LW filter was already in the convergence phase with N=10000.\
In the case of the EuroStoxx50 index, while the Bootstrap filter increases both RMSE and MAE when N is raised to 50000, the APF shows improvements in its performance, according to both criteria, which instead disagree for the Liu and West filter.

@@

Nevertheless, as we previously argued, we should not let tiny differences drive our conclusions. The idea behind employing also such 50000 particles approximations (\textit{i.e} more precise and computationally expensive ones) is to check whether the relative performance that we assessed with N=10000 might be significantly influenced by issues of convergence among the filters: possibly, by further approaching the exact filtering distribution implied by the basic SV model, the filters built on such specification would improve their relative performance. This does not seem to be the case, even though the use of a proxy of the latent volatility process leaves room for discussion, as such conclusions rely on the assumption that such proxy is indeed a good one.\
Rather, what this exercise highlights (under its assumptions) is a need for more flexibility and variability in the basic SV model: the underlying SV model of the LW filter meets this request, but clearly other solutions are possible, and have been argued for in the field of financial econometrics and mathematical finance, as we noted in the literary review section.
